The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-10-17 slides/neon_short_slides.tex
Added short slides

@@

%-------------------------------------------------------------------------------------
\begin{frame}[plain]
  \titlepage
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Moving beyond the spectral norm and Muon}
    \begin{block}{Objective:  $\min_{\mX\in\R^{m\times n}} f(\mX)$}
    \faQuestionCircle \space What will happen if we change the spectral norm in the derivation of Muon?
    \end{block}
    \begin{block}{F-Fanions: Muon, Neon, NSGD, Dion without EF, and so much more}
        From $\norms{\cdot}$ and Muon:
        $$\mX^{t+1} = \mX^{t} - \eta U V^\top$$
        To $\normkfk{\cdot}^\dagger$ and a general F-Fanion:
        $$\mX^{t+1} = \mX^{t} - \eta \left(\alpha \sum_{i=1}^k{u_i v_i^\top} + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\right), \alpha \in [0,1]$$
    \end{block}

\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Linear Minimization Oracle (LMO) and Trust Region}
% after omitting gamma_k, we obtain:
Let us equip $\Rmn$ with a norm $\norm{\cdot}$. Its dual is $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$. $\<\cdot, \cdot>$ is a Frobenius product.
\vspace{0.4em}

Both LMO and Trust Region lead to the update
$$\mX^{t+1} = \mX^{t} - \eta \arg\max_{\mX \in \cB_1} \<\mM^t, \mX> = \mX^{t} - \eta \{\Delta \in \cB_1 \mid \<\mM^t, \Delta> = \norm{\mM^t}^\dagger\}$$

@@

\textbf{Recipe}: It means we seek $\Delta$ from the 1-norm ball that delivers $\<\mM^t, \Delta> = \norm{\mM^t}^\dagger$. We will often return to the SVD $\mM^t = \mU \Sigma \mV^\top$.
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Frobenius $\normf{\mM^{k}}$ and Normalized SGD}

\begin{block}{Deriving NSGD by Recipe} $\normf{\mM^t}^\dagger = \normf{\mM^t}$, and $\Delta = \frac{\mM^t}{\normf{\mM^t}}$ with $\normf{\Delta} = 1$ delivers it. Hence,
$$\mX^{t+1} = \mX^{t} - \eta \frac{\mM^t}{\normf{\mM^t}}$$
\end{block}
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Spectral $\norms{\mM^{k}}$ and Muon. Nuclear $\normn{\mM^k}$ and Neon.}
\begin{block}{Deriving Muon by Recipe} $\norms{\mM^t}^\dagger = \normn{\mM^t}$, and $\Delta = \mU\mV^\top$ delivers it: $\norms{\Delta}$ = 1 and

$$\<\mU\Sigma\mV^\top, \mU\mV^\top> = \tr(\mV \Sigma \mU^\top \mU \mV^\top) = \tr\Sigma = \normn{\mM^t}$$

Hence,
$$\mX^{t+1} = \mX^{t} - \eta \mU \mV^\top$$

\end{block}

\begin{block}{Deriving Neon by Recipe} $\normn{\mM^t}^\dagger = \norms{\mM^t}$, and $\Delta = u_1 v_1^\top$ delivers it: $\normn{\Delta}$ = 1 and

$$\<\mU\Sigma\mV^\top, u_1 v_1^\top> = \tr(\mV \Sigma \mU^\top u_1 v_1^\top) = \tr\diag({\sigma_1, 0, \dots, 0}) = \sigma_1 = \norms{\mM^t}$$

Hence,
$$\mX^{t+1} = \mX^{t} - \eta u_1 v_1^\top$$
\end{block}
\end{frame}

%-------------------------------------------------------------------------------------
\begin{frame}{Of Matrix and Vector Algorithms}
    %\centering
    %\def\arraystretch{1.1}
    \footnotesize
    \begin{table}
    \caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms.}
    \label{tbl:mat_vs_vec_lmo}
    \begin{tabularx}{\linewidth}{|>{\raggedright\arraybackslash}X|c|c|c|}
    \hline
    Method & lmo constraint set $\mathcal D$ & lmo & Reference \\
    \hline\hline
    Normalized SGD & $l_2$-ball, $S_2$-ball & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond} \\
    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
    \hline
    SignSGD & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 1]{bernstein2018signsgd} \\
    Signum  & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 3]{bernstein2018signsgd} \\
    \hdashline
    Muon & Ball in Spectral $S_\infty$ & $-\eta UV^\top$ & \citep{jordan2024muon} \\
    \hline
    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p.19]{shi2016primer}\\
    \hdashline 
    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
    \hline
    \end{tabularx}
    \end{table}
\end{frame}
%------------------------------------------------------------------------------------------------------------
\begin{frame}{Understanding Dion by Thomas Pethick}
    Without momentum, Dion is simplified to
    $$
    \begin{aligned}
    \Delta &\leftarrow g + e  \\
    e &\leftarrow \Delta - \sum_{i=1}^r \sigma_i u_i v_i^\top \\
    x &\leftarrow x - \gamma \sum_{i=1}^r u_i v_i^\top
    \end{aligned}
    $$
    where $\gamma > 0$ and $\sum_{i=1}^r \sigma_i u_iv^\top_i$ is the rank-$r$ truncated SVD of $\Delta \in \Rmn$.
    \vspace{0.4em}

    \faQuestionCircle \space What will we get if we try $\normkfk{\mM}:=\sum_{i=1}^r{\sigma_i}$?
    \end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Ky Fan $k$-rank $\normkfk{\mM^k}^\dagger$ and Fanions}

@@

    Hence,
    $$\mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k} u_i v_i^\top$$
    \end{block}
\end{frame}
%-----------------------------------------------------------------------------------------------------------
\begin{frame}{Computing the $k$-rank update: Lanczos method}
    \begin{center}
    \small
    \begin{tabular}{lccc}
      \toprule
      Method & rtol & $k$ & time (s) \\
      \midrule
      Power Iterations & 0.01 & 1 & 7.7 \\
      SVDS (TRLan) & 0.01 & 1 & \textbf{0.18} \\
      PCA Low Rank (RSVD) & 0.01 & 1 & 1.15 \\
      SVDS (TRLan) & 0.01 & 10 & \textbf{0.47} \\
      PCA Low Rank (RSVD) & 0.01 & 10 & 19.4 \\
      SVDS (TRLan) & 0.01 & 100 & \textbf{1.96} \\
      PCA Low Rank (RSVD) & 0.01 & 100 & 170 \\
      \bottomrule
    \end{tabular}
    \end{center}
    \vspace{0.4em}
    \centering
    \footnotesize Comparison on a \(5000\times5000\) matrix; rtol is the relative Frobenius error of $\sum_{i=1}^{k}u_i \sigma_i v_i^\top$
  \end{frame}
%--------------------------------------------------------------------------------------------------------------
\begin{frame}{Frobeniusize the norms!}
\faExclamationCircle \space $\normf{\cdot}^\dagger = \normf{\cdot}$ is not a Ky Fan norm, so NSGD is not a Fanion.
\vspace{1em}

Let us consider the convex combination of the Ky Fan rank-$k$ norm and the Frobenius norm, which we call F-KF-k-norm: $$\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}, \alpha \in [0,1]$$

@@

        Hence,
        $$\mX^{t+1} = \mX^{t} - \eta \Big(\alpha\sum_{i=1}^{k} u_i v_i^\top + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\Big)$$
        \end{block}
\end{frame}
%------------------------------------------------------------------------------------------------------------
% The next slides were generated by Cursor
\begin{frame}{F-Muon and F-Neon}
  Convex combinations with Frobenius norm:
  \[
    \|\mX\|_{F*}=\alpha\|\mX\|_*+(1-\alpha)\|\mX\|_F,\qquad
    \|\mX\|_{F2}=\alpha\|\mX\|_{op}+(1-\alpha)\|\mX\|_F.
  \]
  Dual-induced updates:
  \[
    \text{F-Muon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,\mU\mV^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big),
  \]
  \[
    \text{F-Neon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,u_1 v_1^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big).
  \]
\end{frame}
%-----------------------------------------------------------------------------------------------
\begin{frame}{Geometric intuition: Dual balls}
    \centering
    \begin{minipage}{0.48\textwidth}
        \includegraphics[height=0.75\textheight,keepaspectratio]{fstardualball.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.48\textwidth}
        \includegraphics[height=0.75\textheight,keepaspectratio]{ftwodualball.pdf}
    \end{minipage}

    \vspace{0.3em}
    \footnotesize Dual balls for F-Muon and F-Neon across \(\alpha\) in a 2D singular-value space
    \end{frame}
%---------------------------------------
\begin{frame}{F-Muon on CIFAR-10 airbench}
    \begin{columns}[T,totalwidth=\textwidth]
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{muon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize With params tuned for Muon, as in \citep{cifar2023airbench}
      \end{column}
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{fmuon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize With params tuned for F-Muon
      \end{column}
    \end{columns}
    \vspace{0.6em}
    \centering

@@

  \end{figure}
\end{frame}
%-----------------------------------------------------------------------------------------------
\begin{frame}{NanoGPT speedrun}
  Setup from \citet{modded_nanogpt_2024}: 1750 iterations and cross-entropy loss lower than 3.28
    \begin{itemize}
      \item Muon: {\tt lr=0.05, momentum=0.95}, final loss = 3.279
      \item F-Muon with \(\alpha=0.5\): {\tt lr=0.07, momentum=0.95}, final loss = 3.281
      \item NSGD: {\tt lr=0.07, momentum=0.96}, final loss = 3.4651
      \item F-Muon with $\alpha=0.5$: {\tt lr=0.07, momentum=0.96}, final loss = 3.2824!
    \end{itemize}
\end{frame}
%----------------------------------------------------------------------------------------------
\begin{frame}{Formal assumptions}
  L-smoothness:
  \begin{aequation}\label{eq:L}
    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
    \quad\text{for all}\;
    \mX, \mX' \in \Rmn,
    \end{aequation}

  Bounded variance:
    \begin{aequation}\label{eq:variance}
      \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
      \quad\text{and}\quad
      \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
      \quad\text{for all}\;
      \mX \in \Rmn,
      \end{aequation}

  Bounded norm:
      \begin{aequation}\label{eq:norm}
        \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
        \quad\text{for all}\;
        \mX \in \Rmn,
        \end{aequation}

  From Marchenko-Pastur law, $\normf{\cdot} \sim \frac{\sqrt{n}}{2}\norms{\cdot}$ and $\normn{\cdot} \sim \frac{n}{2}\norms{\cdot}$ for large $n$.

\end{frame}
\begin{frame}{Bounds from \citep{kovalev2025understanding}}
  \begin{lemma}\label{lemma:stoch_tr}
    To reach the precision $\E{\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger} \leq \epsilon$ under the assumptions \Cref{eq:L}, \Cref{eq:variance}, \Cref{eq:norm}, it is sufficient to choose the parameters as follows:
    \begin{align}
        \eta &= \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
        \qquad
        \alpha = \cO\left(\min\left\{1, \frac{\epsilon^2}{\rho^2\sigma^2}\right\}\right),
        \\
        \label{eq:str_K_nonconvex}
        K &= \cO\left(\max\left\{
            \frac{\rho\sigma}{\epsilon},
            \frac{\rho^3\sigma^3}{\epsilon^3},
            \frac{L\Delta_0}{\epsilon^2},
            \frac{L\Delta_0\rho^2\sigma^2}{\epsilon^4}
        \right\}\right).
    \end{align}
    \end{lemma}

\end{frame}
%----------------------------------------
\begin{frame}{Random linear least squares: Loss and the Spectral Norm of the gradient}
  \begin{columns}[T,totalwidth=\textwidth]
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/loss_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Loss vs iteration
    \end{column}
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/gradient_spectral_norm_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Spectral norm of gradient vs iteration
    \end{column}
  \end{columns}
  \vspace{0.3em}
  \centering
  \footnotesize Linear least squares problem for a 50 x 50 matrix. Common {\tt lr=0.01}.
\end{frame}

%----------------------------------------
\begin{frame}{Geometric intuition: Ky Fan norms}
  \centering
  \begin{minipage}{0.46\textwidth}
      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFan.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.46\textwidth}
      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFanDual.pdf}
  \end{minipage}

  \vspace{0.3em}
  \footnotesize Ky Fan rank-2 ball and its dual in a 3D singular-value space
  \end{frame}
%----------------------------------------
\begin{frame}{Conclusions}
  \begin{itemize}
    \item Muon and NSGD are special cases of F-Fanions
    \item Mixture of norm-based updates is a norm-based update!
    \item No existing bounds describe the superiority of the spectral norm
  \end{itemize}

  \vspace{2em}
  \textbf{For future experiments}: Exploration of matrix and vector algorithms correspondence
\end{frame}

>>>>

% Absolute positioning for overlaying content on the title slide
\usepackage[absolute,overlay]{textpos}

@@

%-------------------------------------------------------------------------------------
\begin{frame}[plain]
  \titlepage
  % Right-side figure on the title slide
  \begin{textblock*}{0.2\paperwidth}(0.75\paperwidth,0.6\paperheight)
    \includegraphics[width=\linewidth]{fstardual_cifar.pdf}
  \end{textblock*}
  \begin{textblock*}{0.22\paperwidth}(0.05\paperwidth,0.55\paperheight)
    \includegraphics[width=\linewidth]{KyFanDual.pdf}
  \end{textblock*}
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Moving beyond the spectral norm and Muon}
    \begin{block}{Objective:  $\min_{\mX\in\R^{m\times n}} f(\mX)$}
    \faQuestionCircle \space What will happen if we change the spectral norm in the derivation of Muon?
    \end{block}
    \begin{block}{F-Fanions: Muon, Neon, NSGD, Dion without EF, and so much more}
        From $\norms{\cdot}$ and Muon:
        $$\mX^{t+1} = \mX^{t} - \eta U V^\top$$
        To $\normkfk{\cdot}^\dagger$ and a general F-Fanion:
        $$\mX^{t+1} = \mX^{t} - \eta \left(\alpha \sum_{i=1}^k{u_i v_i^\top} + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\right), \alpha \in [0,1]$$
    \end{block}

\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Linear Minimization Oracle (LMO)}
% after omitting gamma_k, we obtain:
Let us equip $\Rmn$ with a norm $\norm{\cdot}$. Its dual is $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$. $\<\cdot, \cdot>$ is the Frobenius product.
\vspace{0.4em}

LMO leads to the update
$$\mX^{t+1} = \mX^{t} - \eta \arg\max_{\mX \in \cB_1} \<\mM^t, \mX> = \mX^{t} - \eta \{\Delta \in \cB_1 \mid \<\mM^t, \Delta> = \norm{\mM^t}^\dagger\}$$

@@

\textbf{Recipe}: It means we seek $\Delta$ from the 1-norm ball that delivers $\<\mM^t, \Delta> = \norm{\mM^t}^\dagger$
\vspace{0.4em}

We will often utilize the SVD of $\mM^t$: $\mM^t = \mU \Sigma \mV^\top$.
\end{frame}
\begin{frame}{Of Matrix and Vector Algorithms}
    %\centering
    %\def\arraystretch{1.1}
    \footnotesize
    \begin{table}
    \caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms.}
    \label{tbl:mat_vs_vec_lmo}
    \begin{tabularx}{\linewidth}{|>{\raggedright\arraybackslash}X|c|c|c|}
    \hline
    Method & lmo constraint set $\mathcal D$ & lmo & Reference \\
    \hline\hline
    Normalized SGD & $l_2$-ball, $S_2$-ball & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond} \\
    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
    \hline
    SignSGD & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 1]{bernstein2018signsgd} \\
    Signum  & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 3]{bernstein2018signsgd} \\
    \hdashline
    Muon & Ball in Spectral $S_\infty$ & $-\eta UV^\top$ & \citep{jordan2024muon} \\
    \hline
    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$ & \citep[p.19]{shi2016primer}\\
    \hdashline 
    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
    \hline
    \end{tabularx}
    \end{table}
\end{frame}
%------------------------------------------------------------------------------------------------------------
\begin{frame}{Understanding Dion by Thomas Pethick}
    Without momentum, Dion is simplified to
    $$
    \begin{aligned}
    \Delta &\leftarrow g + e  \\
    e &\leftarrow \Delta - \sum_{i=1}^r \sigma_i u_i v_i^\top \\
    x &\leftarrow x - \gamma \sum_{i=1}^r u_i v_i^\top
    \end{aligned}
    $$
    where $\gamma > 0$ and $\sum_{i=1}^r \sigma_i u_iv^\top_i$ is the rank-$r$ truncated SVD of $\Delta \in \Rmn$.
    \vspace{0.4em}

    \faQuestionCircle \space What will we obtain if we try $\normkfk{\mM}:=\sum_{i=1}^r{\sigma_i}$?
    \end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Ky Fan $k$-rank $\normkfk{\mM^k}^\dagger$ and Fanions}

@@

    Hence,
    $$\mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k} u_i v_i^\top$$
    \end{block}
\end{frame}
%----------------------------------------
\begin{frame}{Geometric intuition: Ky Fan norms}
  \centering
  \begin{minipage}{0.46\textwidth}
      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFan.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.46\textwidth}
      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFanDual.pdf}
  \end{minipage}

  \vspace{0.3em}
  \footnotesize Ky Fan rank-2 ball and its dual in a 3D singular-value space
  \end{frame}
%-----------------------------------------------------------------------------------------------------------
\begin{frame}{Computing the $k$-rank update: Lanczos method}
    \begin{center}
    \small
    \begin{tabular}{lccc}
      \toprule
      Method & rtol & $k$ & time (s) \\
      \midrule
      Power Iterations & 0.01 & 1 & 7.7 \\
      SVDS (TRLan) & 0.01 & 1 & \textbf{0.18} \\
      PCA Low Rank (RSVD) & 0.01 & 1 & 1.15 \\
      SVDS (TRLan) & 0.01 & 10 & \textbf{0.47} \\
      PCA Low Rank (RSVD) & 0.01 & 10 & 19.4 \\
      SVDS (TRLan) & 0.01 & 100 & \textbf{1.96} \\
      PCA Low Rank (RSVD) & 0.01 & 100 & 170 \\
      \bottomrule
    \end{tabular}
    \end{center}
    \vspace{0.4em}
    \centering
    \footnotesize Comparison on a \(5000\times5000\) matrix; rtol is the relative Frobenius error of $\sum_{i=1}^{k}u_i \sigma_i v_i^\top$
  \end{frame}
%--------------------------------------------------------------------------------------------------------------
\begin{frame}{Frobeniusize the norms!}
\faExclamationCircle \space $\normf{\cdot}^\dagger = \normf{\cdot}$ is not a Ky Fan norm, so NSGD is not a Fanion.
\vspace{1em}

Let us consider a convex combination of the Ky Fan rank-$k$ norm and the Frobenius norm, which we call the F-KF-k-norm: $$\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}, \alpha \in [0,1]$$

@@

        Hence,
        $$\mX^{t+1} = \mX^{t} - \eta \Big(\alpha\sum_{i=1}^{k} u_i v_i^\top + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\Big)$$
        \end{block}
\end{frame}
%------------------------------------------------------------------------------------------------------------
% The next slides were generated by Cursor
\begin{frame}{F-Muon and F-Neon}
  Convex combinations with the Frobenius norm:
  \[
    \|\mX\|_{F*}=\alpha\|\mX\|_*+(1-\alpha)\|\mX\|_F,\qquad
    \|\mX\|_{F2}=\alpha\|\mX\|_{op}+(1-\alpha)\|\mX\|_F.
  \]
  Dual-induced updates:
  \[
    \text{F-Muon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,\mU\mV^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big),
  \]
  \[
    \text{F-Neon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,u_1 v_1^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big).
  \]
\end{frame}
%---------------------------------------
\begin{frame}{F-Muon on CIFAR-10 airbench}
    \begin{columns}[T,totalwidth=\textwidth]
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{muon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize With params tuned for Muon, as in \citep{cifar2023airbench}
      \end{column}
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{fmuon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize With params tuned for $\alpha=0.5$ F-Muon
      \end{column}
    \end{columns}
    \vspace{0.6em}
    \centering

@@

  \end{figure}
\end{frame}
%-----------------------------------------------------------------------------------------------
\begin{frame}{NanoGPT speedrun}
  Setup from \citet{modded_nanogpt_2024}: 1750 iterations and cross-entropy loss lower than 3.28
    \begin{itemize}
      \item Muon: {\tt lr=0.05, momentum=0.95}, final loss = 3.279
      \item F-Muon with \(\alpha=0.5\): {\tt lr=0.07, momentum=0.95}, final loss = 3.281
      \item NSGD: {\tt lr=0.07, momentum=0.96}, final loss = 3.4651
      \item F-Muon with $\alpha=0.5$: {\tt lr=0.07, momentum=0.96}, final loss = 3.2824!
    \end{itemize}
\end{frame}
%----------------------------------------
\begin{frame}{Random linear least squares: Loss and the Spectral Norm of the Gradient}
  \begin{columns}[T,totalwidth=\textwidth]
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/loss_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Loss vs iteration
    \end{column}
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/gradient_spectral_norm_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Spectral norm of gradient vs iteration
    \end{column}
  \end{columns}
  \vspace{0.3em}
  \centering
  \footnotesize Linear least squares problem for a 50 x 50 matrix. Common {\tt lr=0.01}.
\end{frame}
%----------------------------------------
\begin{frame}{Conclusion}
  \begin{itemize}
    \item Muon and NSGD are special cases of F-Fanions
    \item A combination of norm-based updates is a norm-based update
    \item No existing bound describes the superiority of the spectral norm
  \end{itemize}

  \vspace{2em}
  \textbf{For future experiments}: Exploration of the matrix and vector algorithms correspondence
\end{frame}
